---
title: "Project 3: Genome Scale algorithms"
author: "Maria, Mateo and Jacob"
date: "2/20/2018"
output:
  pdf_document:
  variant: markdown_github
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggpubr)
library(tidyverse)
```
Instruction of the project: https://github.com/mailund/gsa-exercises/tree/master/Project04

##Project 4 - Suffix arrays and BW-match

## Program description

The programs are implemented in Python and we had no unsolved issues implementing the algorithms.
Execution of the programs on command line takes the form:

```{r, engine = 'python', eval = FALSE, python.reticulate = F}
python search-bs.py [text-filename] [pattern]
python search-bw.py [text-filename] [pattern]
```

The program outputs positions where the second parameter (patern) occurs in the text from file.

##Problems
We did not have any problems regarding implementation except Radix sort. Since sorting suffixes using radix sort requieres creating a 'bucket' for each letter in alphabet and taking ASCII values of letters we could end up with potential problems. Each letter in character coresponds to a integer value called ASCII value but special scharacters could not be found there, which can cause problems when trying to sort them using radix sort. Also implementation of radix sort 


## Evauluation
We measured time for both, search using Burrows-Wheeler transformation and binary search and following plots are showing the results.
```{r}

```


## Test of correctness
We used the test data provided in testdata.zip to test the correctness of the program.

## Time complexity of sorting (python sort and radix sort)

```{r}
bin_search = read.csv("time_binary_search.csv", header = T)
#bin_bwt = read.csv("time_bwt_search.csv", header = T)
radix = read.csv("time_radix.csv", header = T)
sorted = read.csv("time_sorted.csv", header = T)
ggplot(data = bin_search) + 
        geom_point(mapping = aes(y = time/3*log(n), x = n))

ggplot(data = radix) + 
        geom_point(mapping = aes(y = time, x = n))
```



