---
title: "Project 3: Genome Scale algorithms"
author: "Maria, Mateo and Jacob"
date: "2/20/2018"
output:
  pdf_document:
  variant: markdown_github
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggpubr)
library(tidyverse)
```
Instruction of the project: https://github.com/mailund/gsa-exercises/tree/master/Project04

##Project 4 - Suffix arrays and BW-match

## Program description

The program is implemented in Python and we had no unsolved issues implementing the algorithms.
Execution of the program on command line takes the form:

```{r, engine = 'python', eval = FALSE, python.reticulate = F}
python tandem.py [text-filename]
```

The program outputs the number of branching and non-branching tandem repeats to the console, and the full list in a file with same basename as the input file, but with the suffix out.

## Test of correctness
We used the test data provided in testdata.zip to test the correctness of the program.

## Time complexity of sorting (python sort and radix sort)

```{r}
radix = read.csv("time_radix.csv", header = T)
ggplot(data = radix) + 
        geom_point(mapping = aes(y = time/n, x = n)) + theme_bw() 
```



