---
title: "Project 2: Genome Scale algorithms"
author: "Maria, Mateo and Jacob"
date: "2/20/2018"
output:
  pdf_document:
  variant: markdown_github
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggpubr)
```

##Project 2 - Suffix tree construction

## Program description

The programs are implemented in Python, we had no unsolved issues implementing the algorithms.
Execution of the program on commandline takes the form:
  python suffix_tree_naive.py [text-filename] [str-pattern]
and the output is directly to the console.

## Test of search-suffix-tree-naive
We used the data provided for testing which outputs the following results:
python suffix_tree_naive.py mississippi.txt ss
[3, 6]


## Time complexity
```{r}
df_build.naive = read.csv( "Test - SuffixTree Build - Naive.csv" )
colnames(df_build.naive) = c("Function", "Time", "Test", "text.len", "pattern.len", "occurrences", "?")
df_search = read.csv( "Test - SuffixTree Search.csv" )
colnames(df_search) = c("Function", "Time", "Test", "text.len", "pattern.len", "occurrences", "?")

fixed_text_size_build <- df_build.naive %>%
  filter(text.len == 500)

fixed_patter_size_build <- df_build.naive %>%
  filter(pattern.len == 10)

fixed_text_size_search <- df_search %>%
  filter(text.len == 500)

fixed_pattern_size_search <- df_search %>%
  filter(pattern.len == 10)


plot1 <- ggplot(data = fixed_patter_size_build) + 
  geom_point(mapping = aes(x = text.len/Time^2, y = Time)) +
  ggtitle("Building tree: \nFixed pattern size: n^2 time") +
  geom_hline(yintercept = 0.01, color = "red")

plot2 <- ggplot(data = fixed_text_size_build) + 
  geom_point(mapping = aes(x = pattern.len/Time^2, y = Time)) + 
  geom_hline(yintercept = 0.18, color = "red") +
  ggtitle("Building tree: \nFixed text size: n^2 time")

plot3 <- ggplot(data = fixed_text_size_search) + 
  geom_point(mapping = aes(x = pattern.len/Time^2, y = Time)) + 
  geom_hline(yintercept = 0.00008, color = "red") +
  ggtitle("Searching tree: \nFixed text size: n^2 time")

#we dont need fourth plot since, searching the tree is not affected by text size, only the pattern

ggarrange(plot1,plot2,plot3,
      ncol = 2, nrow =2) 
  
```

```{r}

ggplot(data = df_build.naive) + 
  geom_point(mapping = aes(x = pattern.len, y = Time, color = Test)) +
  ggtitle("")

```


Instruction of the project: https://github.com/mailund/gsa-exercises/tree/master/Project02

